---
title: "ConnectionContextDemo.Rmd"
author: "R ML library team"
date: "9/23/2019"
output: html_document
---

## Connection Context Example

HANA ML R API supports both JDBC and ODBC connection to HANA DB:

```{r}
library(hana.ml.r)
```

## JDBC Connection Context

To use JDBC, you need to check whether JDK and rJava has been installed or not. Use option jdbcDriverPath to point to HANA client jdbc driver (ngdbc.jar) and set odbc = FALSE:

```{r}
conn <- hanaml.ConnectionContext(dsn = 'xxx.xxx.xxx.xxx:30015',
                                 user = 'DM_PAL',
                                 password = 'xxxxxxxxx',
                                 odbc = FALSE,
                                 jdbcDriver = "/home/machine/sap/hdbclient/ngdbc.jar")
```

## ODBC Connection Context
To use ODBC, you need to set up HANA instance in the odbc driver manager.
```{r}
conn2 <- hanaml.ConnectionContext(dsn = 'DEV',
                                 user = 'DM_PAL',
                                 password = 'Abcd1234')
```


## GetCurrentSchema()
Return the current schema.

```{r}
conn$GetCurrentSchema()
```

## Create HANA DataFrame referencing to HANA table
table() provides a way to create HANA DataFrame. 
```{r}
hana.df <- conn$table("IRIS_DATA_FULL_TBL")
```
```{r}
hana.df$Head(5)$Collect()
```
