---
title: "DataFrameDemo"
author: "PAL team"
date: "9/25/2019"
output: html_document
---

## HANA DataFrame
HANA DataFrame is composed of HANA ConnectionContext and SQL statement. It represents a database query as a DataFrame. Most operations of HANA DataFrame are designed not bring data back from the database unless explicityly requested.

```{r}
conn <- hanaml.ConnectionContext(dsn = 'xxx.xxx.xxx.xxx:30015',
                                 user = 'DM_PAL',
                                 password = 'xxxxxxxx',
                                 odbc = FALSE,
                                 jdbcDriver = "/home/machine/sap/hdbclient/ngdbc.jar",
                                 identifier.quote = "\"")
```

```{r}

library(datasets)
data(iris)
summary(iris)
```

```{r}
hana.df <- ConvertToHANADataFrame(conn, iris, "IRIS_DATA", force = TRUE, native = TRUE)
```

```{r}
hana.df <- conn$table("IRIS_DATA")
```

## Collect()
Collect() performs a sqlQuery to fetch the data from HANA DB, it can be used together with Head(N) to fetch top N rows:

```{r}
hana.df$Head(10)$Collect()
```

## Count()
Count performs a sqlQuery to fetch the row number of the data from HANA DB.

```{r}
hana.df$Count()
```

## Sort(cols)
Sort the data in HANA DB according to a list of columns.

```{r}
hana.df$Sort(list("Sepal.Length"))$Head(10)$Collect()
```

## Drop(cols)
Create a new select statement to remove a list of columns.

```{r}
hana.df$Drop(list("Sepal.Length"))$Head(10)$Collect()
```

## AddId()
Create ID columns.
```{r}
hana.df$AddId("NEW_ID")$Head(10)$Collect()
```

## DropDuplicates()
Remove duplicated rows.
```{r}
hana.df$DropDuplicates()$Head(10)$Collect()
```

## Join(other)
Join two HANA DataFrame.
```{r}
hana.df$Alias("L")$Join(hana.df$Alias("R"), on_expression = "L.\"Sepal.Length\" = R.\"Sepal.Length\"")$Head(10)$Collect()
```

## Select(cols)
Create a new select statement to select subset of columns.
```{r}
hana.df$Select("Sepal.Length")$Head(10)$Collect()
```

## save(table)
Persist data to a new HANA table.
```{r}
hana.df$save(table = "TEST", force = TRUE, schema = "DM_PAL")
```

## dtypes()
Return table structure of HANA table.
```{r}
hana.df$dtypes(list("Sepal.Length", "Species"))
```

```{r}
hana.df$columns.df
```

## Describe()
Return statistics of HANA table.
```{r}
hana.df$Describe()
```

## DataManipulation()
Unload table from memory.
```{r}
DataManipulation(conn, "IRIS_DATA")
```

